Looking for a Signal in the Noise: Revisiting Obesity and the Microbiome

### 2016-06-03 submission/supp_text.Rmd
Incorporate supplementary text file

---
title: "Looking for a Signal in the Noise: Revisiting Obesity and the Microbiome"
output:
  html_document:
    keep_md: yes
bibliography: reference.bib
---

Marc A Sze<sup>1</sup> and Patrick D Schloss<sup>1</sup>

```{r Loading_data_tables, echo=FALSE, message=FALSE, warning=FALSE}

alpha_tests.summary <- read.table("../data/process/alpha_tests.summary", header=T)
beta_tests.summary <- read.table("../data/process/beta_tests.summary", header=T)
```

###In-Depth Overview of Search Strategy:

@@

*Inclusion Criteria:*

@@

*Exclusion Criteria:*

@@

###Supplemental Results:

```{r load_needed_data_tables, echo=FALSE, message=FALSE, warning=FALSE}

alpha_tests.summary <- read.table("../data/process/alpha_tests.summary", header=T)
alpha_composite.summary <- read.table("../data/process/alpha_composite.summary")

```

**Figure S1: Individual and Combined comparison of Obese and Non-Obese groups for Observed OTUs.**  For observed OTUs Baxter (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'baxter' & alpha_tests.summary$metric == 'sobs'), 'p.value'], 2), nsmall = 2)`, Goodrich (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'goodrich' & alpha_tests.summary$metric == 'sobs'), 'p.value'], 8), nsmall = 8)`), and Turnbaugh (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'turnbaugh' & alpha_tests.summary$metric == 'sobs'), 'p.value'], 2), nsmall = 2)`) studies individually were signficantly different for obese and non-obese groups.  When the data sets were combined and study effect was corrected for,  there was a significant difference between Non-Obese and Obese groups (P-value = `r format(round(alpha_composite.summary['sobs', 'composite_p'], 8), nsmall = 8)`)

**Figure S2: Individual and Combined comparison of Obese and Non-Obese groups for Evenness.**  For evenness only the Goodrich study (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'goodrich' & alpha_tests.summary$metric == 'shannoneven'), 'p.value'], 3), nsmall = 3)`)  individually was signficantly different for obese and non-obese groups.  When the data sets were combined and study effect was corrected for,  there was an overall significant difference between Non-Obese and Obese groups (P-value = `r format(round(alpha_composite.summary['shannoneven', 'composite_p'], 2), nsmall = 2)`)

**Figure S3: Individual and Combined comparison of Obese and Non-Obese groups for Bacteroidetes Relative Abundance.**  Only the Escobar study (P-value = `r format(round(alpha_tests.summary[which(alpha_tests.summary$dataset == 'escobar' & alpha_tests.summary$metric == 'bacteroidetes'), 'p.value'], 3), nsmall = 3)`)  individually was signficantly different for obese and non-obese groups.  When the data sets were combined and study effect was corrected for,  there was no overall significant difference between Non-Obese and Obese groups (P-value = `r format(round(alpha_composite.summary['bacteroidetes', 'composite_p'], 2), nsmall = 2)`)

**Figure S4: Individual and Combined comparison of Obese and Non-Obese groups for Firmicutes Relative Abundance.**  No study was individually signficantly different for obese and non-obese groups.  When the data sets were combined and study effect was corrected for,  there was also no overall significant difference between Non-Obese and Obese groups (P-value = `r format(round(alpha_composite.summary['firmicutes', 'composite_p'], 2), nsmall = 2)`)

**Figure S5: Funnel Plot of the Shannon Diversity and Bacteroidetes/Firmicutes Ratio Relative Risk.**  **A)** Overall there does not seem to be any bias associated with the studies selected for the Shannon Diversity analysis with studies falling on either side of the predicted value and those with smaller total n falling further away from this. **B)** For the Bacteroidetes/Firmicutes ratio there does not seem to be any associated bias. The overall pattern is similar to that observed for the Shannon Diversity analysis.

**Figure S6. Summary of Power and Sample Size Simulations for OTU Richness (S) for Non-obese versus Obese**  P-values listed in the legend represent the outcome of a wilcoxson rank sum test between non-obese and obese for each specific data set.  **A)** The dotted lines represent the actual effect size for each specific study.  **B)** The dotted lines represent the needed n to achieve an 80% power with the actual study effect size.

**Figure S7. Summary of Power and Sample Size Simulations for Evenness (J) for Non-obese versus Obese**  P-values listed in the legend represent the outcome of a wilcoxson rank sum test between non-obese and obese for each specific data set.  **A)** The dotted lines represent the actual effect size for each specific study.  **B)** The dotted lines represent the needed n to achieve an 80% power with the actual study effect size.

**Figure S8. Summary of Power and Sample Size Simulations for Bacteroidetes for Non-obese versus Obese**  P-values listed in the legend represent the outcome of a wilcoxson rank sum test between non-obese and obese for each specific data set.  **A)** The dotted lines represent the actual effect size for each specific study.  **B)** The dotted lines represent the needed n to achieve an 80% power with the actual study effect size.

**Figure S9. Summary of Power and Sample Size Simulations for Firmicutes for Non-obese versus Obese**  P-values listed in the legend represent the outcome of a wilcoxson rank sum test between non-obese and obese for each specific data set.  **A)** The dotted lines represent the actual effect size for each specific study.  **B)** The dotted lines represent the needed n to achieve an 80% power with the actual study effect size.

**Figure S10. Summary of Power and Sample Size Simulations for Relative Risk  for Obesity based on Shannon Diversity**  P-values listed in the legend represent the outcome of a wilcoxson rank sum test between non-obese and obese for each specific data set.  **A)** The dotted lines represent the actual effect size for each specific study.  **B)** The dotted lines represent the needed n to achieve an 80% power with the actual study effect size.

**Figure S11. Summary of Power and Sample Size Simulations for Relative Risk  for Obesity based on B/F Ratio**  P-values listed in the legend represent the outcome of a wilcoxson rank sum test between non-obese and obese for each specific data set.  **A)** The dotted lines represent the actual effect size for each specific study.  **B)** The dotted lines represent the needed n to achieve an 80% power with the actual study effect size.

*******
##### References:

>>>>

---
title: ""
bibliography: reference.bib
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header.tex
csl: mbio.csl
fontsize: 12pt
geometry: margin=1.0in
---

## Supplementary Text 1: Looking for a Signal in the Noise: Revisiting Obesity and the Microbiome

Marc A Sze and Patrick D Schloss

### In-Depth Overview of Search Strategy

@@

***Inclusion Criteria:***

@@

***Exclusion Criteria:***

@@

### References
